---
title: "startr"
output: 
  flexdashboard::flex_dashboard:
    navbar:
     - { title: "Developed by PainblogR", icon: 'fa-home', href: "https://www.painblogr.org", align: right }
    social: ['menu']
    source_code: 'https://github.com/kamermanpr/painblogr-startr.git'
runtime: shiny
---

```{r setup, include = FALSE}
library(flexdashboard)
library(tidyverse)
library(lme4)
library(skimr)
library(readxl)
library(car)
library(gmodels)

skim_with(numeric = list(p0 = NULL, p25 = NULL, p50 = NULL, 
                         p75 = NULL, p100 = NULL, hist = NULL,
                         min = ~ min(., na.rm = TRUE),
                         q25 = ~ quantile(., probs = 0.25, na.rm = TRUE),
                         median = ~ quantile(., probs = 0.5, na.rm = TRUE),
                         q75 = ~ quantile(., probs = 0.75, na.rm = TRUE),
                         max = ~min(., na.rm = TRUE)))
```

Welcome {data-navmenu='Select a test'}
===================================== 

Welcome to the `startr` app.

**To start, choose the type of test using the menu labeled _"Select a test"_.** 

The app provides five groups of statistical tests that typically are taught in introductory biostatistcs courses. These five groups, and the tests they include, are:

1. Comparisons between two groups  
    - Unpaired continuous data: Student's t-test and Mann-Whitney U test  
    - Paired continuous data: Student's t-test and Wilcoxon signed-rank test  

2. Comparisons between $\geq$ 3 groups  
    - One-way Analysis of Variance (ANOVA) and Kruskal-Wallis test  
    - Repeated-measures one-way ANOVA and Friedman test  

3. Correlation  
    - Pearson's product-moment correlation  
    - Spearman's rank-order correlation  

4. Linear regression  

5. Contingency tables  
    - Chi-squared tests ($\geq$ 2x2 tables)  
    - Fisher's Exact test ($\geq$ 2x2 tables)  
    - Mcnemar's test (repeated measures designs, 2x2 tables only)

On the left side of each page associated with a group of tests (accessed through the _"Select a test"_ menu), detailed instructions are provided on how to enter the data. 

**Please note:** In all cases where the test is an omnibus test (e.g., ANOVA), no pairwise _post-hoc_ tests are provided. 
    
Compare 2 groups {data-navmenu='Select a test'}
=====================================  

Inputs {.sidebar data-width=350}
-------------------------------------

#### Two groups

##### PROCEDURE
1. Select a hypothesis test (default value: paired Student's _t_-test).  
2. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV).  
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it.  

##### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV/XLS/XLSX file consisting of two columns.  
- For paired analyses, there must be one column for each of the two measurements made on an animal/person (incomplete pairs will be removed).  
- For unpaired tests, there must be one column for each of the two groups data were collected from.  
- See examples of the data structure  <a href="https://drive.google.com/file/d/1dwelSQU61MRZALmlwLJ_IuOsy8OBHJY9/view?usp=sharing" target="_blank">here</a>

```{r}
hr()
selectInput(inputId = 'test0', 
            label = 'Choose a test',
            choices = c('Paired: t-test', 'Paired: Wilcoxon signed-rank test',
                        'Unpaired: t-test', 'Unpaired: Mann-Whitney U test'))

checkboxInput(inputId = 'type0', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header0', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file0', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
df <- reactive({
    req(input$file0)
    
    if(input$type0 == TRUE) {
        read_csv(input$file0$datapath,
             col_names = input$header0)
    } else {
        read_excel(input$file0$datapath,
                   col_names = input$header0)
    }
})

renderPrint({
    df()
})
```

### Tabular summary

```{r}
renderPrint({
    skim(df())
})
``` 

Row
-------------------------------------  

### Box-and-whisker plot

```{r}
renderPlot({
    gathered <- df() %>% 
        gather(key = key,
               value = value)
    
    ggplot(data = gathered) +
        aes(x = key,
            y = value) +
    geom_point(position = position_jitter(width = 0.3)) +
    geom_boxplot(alpha = 0.5) +
    stat_summary(fun.y = 'mean', 
                 geom = 'point',
                 shape = 8,
                 stroke = 1,
                 size = 4) +
    scale_y_continuous(name = 'Value') +
    theme_bw(base_size = 18) +
    theme(axis.title.x = element_blank())
    })
```

### Test summary

```{r}
renderPrint({
    gathered <- df() %>% 
        gather(key = key,
               value = value)
    
    gathered_paired <- df() %>% 
        filter(complete.cases(.)) %>% 
        gather(key = key,
               value = value)
    
    if(input$test0 == 'Paired: t-test') {
        t.test(value ~ key,
               data = gathered_paired,
               paired = TRUE)
        
    } else if(input$test0 == 'Unpaired: t-test') {
        t.test(value ~ key,
               data = gathered,
               na.rm = TRUE,
               paired = FALSE)
        
    } else if(input$test0 == 'Paired: Wilcoxon signed-rank test') {
        wilcox.test(value ~ key, 
                    data = gathered_paired,
                    exact = FALSE,
                    paired = TRUE,
                    confint = TRUE)
        
    } else {
        wilcox.test(value ~ key,
                    data = gathered,
                    na.rm = TRUE,
                    exact = FALSE,
                    paired = FALSE,
                    confint = TRUE)
    }
})
```

Compare $\geq$ 3 groups {data-navmenu='Select a test'}
=====================================  

Inputs {.sidebar data-width=350}
-------------------------------------

#### Analysis of Variance

##### PROCEDURE
1. Select a hypothesis test (default value: ANOVA).  
2. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV).  
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it.  

##### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV/XLS/XLSX file consisting of two columns.  
- The first column for repeated measures designs (RM-ANOVA and Friedman test) is assumed to be the column that contains the grouping variable (e.g., subject ID).  
- Missing data pairs will be removed for repreated measures designs. 
- See examples of the data structure <a href="https://drive.google.com/file/d/1D-cNaNxQnbbkWbXTdRRgvl0W1E7XMO5l/view?usp=sharing" target="_blank">here</a>

```{r}
hr()
selectInput(inputId = 'test1', 
            label = 'Choose a test',
            choices = c("ANOVA", "Repeated-measures ANOVA",
                        "Kruskal-Wallis test", "Friedman test"))

checkboxInput(inputId = 'type1', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header1', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file1', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
aov_data <- reactive({
    req(input$file1)
    
    if(input$type1 == TRUE) {
        read_csv(input$file1$datapath,
                 col_names = input$header1)
    } else {
        read_excel(input$file1$datapath,
                   col_names = input$header1)
    }
})

renderPrint({
    aov_data()
})
```

### Tabular summary

```{r}
renderPrint({
    if(input$test1 == 'ANOVA' | input$test1 == 'Kruskal-Wallis test'){
        skim(aov_data())
    } else {
        nam_id <- as.name(names(aov_data()[1]))
        
        aov_data() %>% 
            mutate(ID = as.character(!!nam_id)) %>% 
            skim()
    }
})
``` 

Row
-------------------------------------  

### Scatterplot

```{r}
renderPlot({
    if(input$test1 == 'ANOVA' | input$test1 == 'Kruskal-Wallis test') {
        gathered <- aov_data() %>% 
            gather(key = group,
                   value = value)
        
        ggplot(data = gathered) +
            aes(x = group,
                y = value) +
            geom_point(position = position_jitter(width = 0.3)) +
            geom_boxplot(alpha = 0.5) +
            stat_summary(fun.y = 'mean', 
                         geom = 'point',
                         shape = 8,
                         stroke = 1,
                         size = 4) +
            scale_y_continuous(name = 'Value') +
            theme_bw(base_size = 18) +
            theme(axis.title.x = element_blank())
    } else {
        nam_id <- as.name(names(aov_data()[1]))
        
        fig_rep <- aov_data() %>% 
            gather(key = group,
                   value = value,
                   - !!nam_id)
        
        ggplot(data = fig_rep) +
            aes(x = group,
                y = value) +
            geom_point(position = position_jitter(width = 0.3)) +
            geom_boxplot(alpha = 0.5) +
            stat_summary(fun.y = 'mean', 
                         geom = 'point',
                         shape = 8,
                         stroke = 1,
                         size = 4) +
            scale_y_continuous(name = 'Value') +
            theme_bw(base_size = 18) +
            theme(axis.title.x = element_blank())
    }
    
})
```

### ANOVA summary

```{r}
renderPrint({
    unpaired <- aov_data() %>% 
        gather(key = group,
               value = value)
    
    nam_id <- as.name(names(aov_data()[1]))
    
    paired <- aov_data() %>% 
        filter(complete.cases(.)) %>%
        gather(key = group,
               value = value,
               -!!nam_id) %>% 
        rename(ID = !!nam_id) %>% 
        mutate(ID = as.character(ID))
    
    if(input$test1 == 'ANOVA') {
        Anova(lm(value ~ group,
                 data = unpaired),
              test.statistic = 'F')
        
    } else if(input$test1 == 'Kruskal-Wallis test') {
        kruskal.test(value ~ factor(group),
                     data = unpaired)
        
    } else if(input$test1 == 'Repeated-measures ANOVA') {
        Anova(lmer(value ~ group + (1 | ID), 
                   data = paired),
              test.statistic = 'F')
    } else {
        friedman.test(value ~ group | ID, 
                      data = paired)
    }
})
```

Correlation {data-navmenu='Select a test'}
=====================================  
    
Inputs {.sidebar data-width=350}
-------------------------------------

#### Correlation

##### PROCEDURE
1. Select the type of correlation (default value: Pearson's correlation). 
2. Select whether you want a trend line included on the scatterplot (default value: true).
3. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV). 
4. Select whether the first row of the uploaded file includes column headings (default value: true).  
5. Select your data file and upload it. 

##### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV/XLS/XLSX file consisting of two columns.  
- Missing data pairs will be removed.  
- See examples of the data structure  <a href="https://drive.google.com/file/d/1PeWH4De1TKi_yfhGMyhJxpqIu331RhEq/view?usp=sharing" target="_blank">here</a>

```{r}
hr()
selectInput(inputId = 'test2', 
            label = 'Choose a test',
            choices = c("Pearson's correlation", "Spearman's correlation"))

checkboxInput(inputId = 'trend_line2', 
              label = 'Draw a trend line.', 
              value = TRUE)

checkboxInput(inputId = 'type2', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header2', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file2', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
cor_data <- reactive({
    req(input$file2)
    
    if(input$type2 == TRUE) {
        read_csv(input$file2$datapath,
                 col_names = input$header2)
    } else {
        read_excel(input$file2$datapath,
                   col_names = input$header2)
    }
})

renderPrint({
    cor_data()
})
```

### Tabular summary

```{r}
renderPrint({
    skim(cor_data())
})
``` 

Row
-------------------------------------  

### Scatterplot

```{r}
renderPlot({
    if(input$trend_line2 == TRUE) {
        ggplot(data = cor_data()) +
            aes_string(x = colnames(cor_data())[1],
                       y = colnames(cor_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            geom_smooth(method = 'loess',
                        se = FALSE) +
            theme_bw(base_size = 18)
    
    } else {
        ggplot(data = cor_data()) +
            aes_string(x = colnames(cor_data())[1],
                       y = colnames(cor_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            theme_bw(base_size = 18)
    }
})
```

### Correlation summary

```{r}
renderPrint({
    
    corr <- cor_data() %>% 
        filter(complete.cases(.)) %>% 
        set_names(c('column_1.y', 'column_2.x'))
    
    if(input$test2 == "Pearson's correlation") {
        with(corr, cor.test(x = column_2.x, 
                            y = column_1.y, 
                            method = 'pearson'))
    } else {
        with(corr, cor.test(x = column_2.x, 
                            y = column_1.y, 
                            method = 'spearman',
                            exact = TRUE))
    }
})
```

Linear regression {data-navmenu='Select a test'}
=====================================  
    
Inputs {.sidebar data-width=350}
-------------------------------------

#### Linear regression 

##### PROCEDURE
1. Select whether you want to plot 95% confidence intervals for the regression line (default value: true). 
2. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV). 
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it. 

##### NOTE 
- Ordinary least squares regression (OLS) only.  
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV/XLS/XLSX file consisting of two columns.  
- The first column is assumed to be the dependent variable (_y_).  
- Missing data pairs will be removed.  
- See examples of the data structure  <a href="https://drive.google.com/file/d/1FPfJnQ_ZCTMSe0SVtO2RhagpL4lmWe__/view?usp=sharing" target="_blank">here</a> 


```{r}
hr()
checkboxInput(inputId = 'confint3', 
              label = 'Plot 95% confidence interval.', 
              value = TRUE)

checkboxInput(inputId = 'type3', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header3', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file3', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
lm_data <- reactive({
    req(input$file3)
    
    if(input$type3 == TRUE) {
        read_csv(input$file3$datapath,
                 col_names = input$header3)
    } else {
        read_excel(input$file3$datapath,
                   col_names = input$header3)
    }
})

renderPrint({
    lm_data()
})
```

### Tabular summary

```{r}
renderPrint({
    skim(lm_data())
})
``` 

Row
-------------------------------------  

### Scatterplot

```{r}
renderPlot({
    if(input$confint3 == TRUE) {
        ggplot(data = lm_data()) +
            aes_string(x = colnames(lm_data())[2],
                       y = colnames(lm_data())[1]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            geom_smooth(method = 'lm') +
            theme_bw(base_size = 18)
    
    } else {
        ggplot(data = lm_data()) +
            aes_string(x = colnames(lm_data())[2],
                       y = colnames(lm_data())[1]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
             geom_smooth(method = 'lm',
                         se = FALSE) +
            theme_bw(base_size = 18)
    }
})
```

### Regression summary

```{r}
renderPrint({
    ols <- lm_data() %>% 
        filter(complete.cases(.)) %>% 
        set_names(c('column_1.y', 'column_2.x'))
    
    summary(lm(column_1.y ~ column_2.x, data = ols))
})
```

Contingency tables {data-navmenu='Select a test'}
=====================================  
    
Inputs {.sidebar data-width=350}
-------------------------------------

#### Contingency tables

##### PROCEDURE
1. Select whether you are entering raw data or summarised counts (default value: raw data). 
2. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV). 
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it. 

##### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload count data only.  
- If uploading raw data, the file must be a CSV/XLS/XLSX file consisting of two columns, with the data entered in long format.  
- If uploading summarised data, the file must be a CSV/XLS/XLSX file consisting of three columns (two grouping columns and a count column).
- Missing data pairs will be removed.  
- See examples of the data structure  <a href="https://drive.google.com/file/d/1NnQYJgCZridAm9NySPrco9353gjhpkKL/view?usp=sharing" target="_blank">here</a>

```{r}
hr()
selectInput(inputId = 'test4', 
            label = 'Choose a test',
            choices = c("Chi-squared", "Fisher's Exact", 
                        "Mcnemar's"))

selectInput(inputId = 'data_format4', 
            label = 'Choose the data format',
            choices = c("Raw data", "Summarised counts"))

checkboxInput(inputId = 'type4', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header4', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file4', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Column
------------------------------------

### Mosaic plot

```{r}
conti_data <- reactive({
    req(input$file4)
    
    if(input$type4 == TRUE) {
        read_csv(input$file4$datapath,
                 col_names = input$header3)
    } else {
        read_excel(input$file4$datapath,
                   col_names = input$header3)
    }
})

renderPlot({
    if(input$data_format4 == 'Raw data') {
        par(cex = 2)
        mosaicplot(table(conti_data()),
                   main = 'Mosaic plot of uploaded data',
                   color = TRUE,
                   cex.axis = 0.75)
    } else {
        nam <- as.name(names(conti_data())[[3]])
        
        tab4 <- conti_data() %>% 
            uncount(!!nam)
        
        par(cex = 2)
        
        mosaicplot(table(tab4),
                   main = 'Mosaic plot of uploaded data',
                   color = TRUE,
                   cex.axis = 0.75)
    }
})
```

Column
------------------------------------

### Tabulated data and test results

```{r}
renderPrint({
if(input$test4 == 'Chi-squared') {
    if(input$data_format4 == 'Raw data') {
        CrossTable(x = conti_data()[[1]], 
                   y = conti_data()[[2]],
                   prop.chisq = FALSE,
                   chi = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
    } else {
        nam <- as.name(names(conti_data())[[3]])
        
        tab4 <- conti_data() %>% 
            uncount(!!nam)
        
        CrossTable(x = tab4[[1]], 
                   y = tab4[[2]],
                   prop.chisq = FALSE,
                   chi = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
        }
    } else if(input$test4 == "Fisher's Exact") {
    if(input$data_format4 == 'Raw data') {
        CrossTable(x = conti_data()[[1]], 
                   y = conti_data()[[2]],
                   prop.chisq = FALSE,
                   fisher = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
    } else {
        tab4 <- conti_data() %>% 
            uncount(count)
        
        CrossTable(x = tab4[[1]], 
                   y = tab4[[2]],
                   prop.chisq = FALSE,
                   fisher = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
        }
    } else {
    if(input$data_format4 == 'Raw data') {
        CrossTable(x = conti_data()[[1]], 
                   y = conti_data()[[2]],
                   prop.chisq = FALSE,
                   mcnemar = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
    } else {
        tab4 <- conti_data() %>% 
            uncount(count)
        
        CrossTable(x = tab4[[1]], 
                   y = tab4[[2]],
                   prop.chisq = FALSE,
                   mcnemar = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
        }
    }
})
```
