---
title: "startr"
output: 
  flexdashboard::flex_dashboard:
    navbar:
     - { title: "Developed by PainblogR", icon: 'fa-home', href: "https://www.painblogr.org", align: right }
    social: ['menu']
    source_code: 'https://github.com/kamermanpr/painblogr-startr.git'
runtime: shiny
---

```{r setup, include = FALSE}
library(flexdashboard)
library(tidyverse)
library(broom)
library(skimr)
library(knitr)
library(kableExtra)
library(readxl)
library(car)
library(gmodels)

skim_with(numeric = list(p0 = NULL, p25 = NULL, p50 = NULL, 
                         p75 = NULL, p100 = NULL, hist = NULL,
                         min = ~ min(., na.rm = TRUE),
                         q25 = ~ quantile(., probs = 0.25, na.rm = TRUE),
                         median = ~ quantile(., probs = 0.5, na.rm = TRUE),
                         q75 = ~ quantile(., probs = 0.75, na.rm = TRUE),
                         max = ~min(., na.rm = TRUE)))
```

Two groups (paired and unpaired) {data-navmenu='Select a test'}
=====================================  


Inputs {.sidebar data-width=350}
-------------------------------------

#### Two groups

##### PROCEDURE
1. Select a hypothesis test (default value: paired Student's _t_-test).  
2. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV).  
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it.  

##### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV/XLS/XLSX file consisting of two columns.  
- For paired analyses, there must be one column for each of the two measurements made on an animal/person (incomplete pairs will be removed).  
- For unpaired tests, there must be one column for each of the two groups data were collected from.  

```{r}
hr()
selectInput(inputId = 'test', 
            label = 'Choose a test',
            choices = c('Paired: t-test', 'Paired: Wilcoxon signed-rank test', 
                        'Unpaired: t-test', 'Unpaired: Mann-Whitney U test'))

checkboxInput(inputId = 'type', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
df <- reactive({
    req(input$file)
    
    if(input$type == TRUE) {
        read_csv(input$file$datapath,
             col_names = input$header)
    } else {
        read_excel(input$file$datapath,
                   col_names = input$header)
    }
})

renderPrint({
    df()
})
```

### Tabular summary

```{r}
renderPrint({
    skim(df())
})
``` 

Row
-------------------------------------  

### Box-and-whisker plot

```{r}
renderPlot({
    gathered <- df() %>% 
        gather(key = key,
               value = value)
    
    ggplot(data = gathered) +
        aes(x = key,
            y = value) +
    geom_point(position = position_jitter(width = 0.3)) +
    geom_boxplot(alpha = 0.5) +
    stat_summary(fun.y = 'mean', 
                 geom = 'point',
                 shape = 8,
                 stroke = 1,
                 size = 4) +
    scale_y_continuous(name = 'Value') +
    theme_bw(base_size = 20) +
    theme(axis.title.x = element_blank())
    })
```

### Test summary

```{r}
renderPrint({
    gathered <- df() %>% 
        gather(key = key,
               value = value)
    
    gathered_paired <- df() %>% 
        filter(complete.cases(.)) %>% 
        gather(key = key,
               value = value)
    
    if(input$test == 'Paired: t-test') {
        t.test(value ~ key,
               data = gathered_paired,
               paired = TRUE)
        
    } else if(input$test == 'Unpaired: t-test') {
        t.test(value ~ key,
               data = gathered,
               na.rm = TRUE,
               paired = FALSE)
        
    } else if(input$test == 'Paired: Wilcoxon signed-rank test') {
        wilcox.test(value ~ key, 
                    data = gathered_paired,
                    exact = FALSE,
                    paired = TRUE,
                    confint = TRUE)
        
    } else {
        wilcox.test(value ~ key,
                    data = gathered,
                    na.rm = TRUE,
                    exact = FALSE,
                    paired = FALSE,
                    confint = TRUE)
    }
})
```

Correlation {data-navmenu='Select a test'}
=====================================  
    
Inputs {.sidebar data-width=350}
-------------------------------------

#### Correlation

##### PROCEDURE
1. Select the type of correlation (default value: Pearson's correlation). 
2. Select whether you want a trend line included on the scatterplot (default value: true).
3. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV). 
4. Select whether the first row of the uploaded file includes column headings (default value: true).  
5. Select your data file and upload it. 

##### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV/XLS/XLSX file consisting of two columns.  
- Missing data pairs will be removed. 

```{r}
hr()
selectInput(inputId = 'test2', 
            label = 'Choose a test',
            choices = c("Pearson's correlation", "Spearman's correlation"))

checkboxInput(inputId = 'trend_line2', 
              label = 'Draw a trend line.', 
              value = TRUE)

checkboxInput(inputId = 'type2', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header2', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file2', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
cor_data <- reactive({
    req(input$file2)
    
    if(input$type2 == TRUE) {
        read_csv(input$file2$datapath,
                 col_names = input$header2)
    } else {
        read_excel(input$file2$datapath,
                   col_names = input$header2)
    }
})

renderPrint({
    cor_data()
})
```

### Tabular summary

```{r}
renderPrint({
    skim(cor_data())
})
``` 

Row
-------------------------------------  

### Scatterplot

```{r}
renderPlot({
    if(input$trend_line2 == TRUE) {
        ggplot(data = cor_data()) +
        aes_string(x = colnames(cor_data())[1],
                   y = colnames(cor_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            geom_smooth(method = 'loess',
                        se = FALSE) +
            theme_bw(base_size = 20)
    
    } else {
        ggplot(data = cor_data()) +
        aes_string(x = colnames(cor_data())[1],
                   y = colnames(cor_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            theme_bw(base_size = 20)
    }
})
```

### Correlation summary

```{r}
renderPrint({
    
    corr <- cor_data() %>% 
        filter(complete.cases(.)) %>% 
        set_names(c('column_1.y', 'column_2.x'))
    
    if(input$test2 == "Pearson's correlation") {
        with(corr, cor.test(x = column_2.x, 
                            y = column_1.y, 
                            method = 'pearson'))
    } else {
        with(corr, cor.test(x = column_2.x, 
                            y = column_1.y, 
                            method = 'spearman',
                            exact = TRUE))
    }
})
```

Regression {data-navmenu='Select a test'}
=====================================  
    
Inputs {.sidebar data-width=350}
-------------------------------------

#### Regression 

##### PROCEDURE
1. Select whether you want to plot 95% confidence intervals for the regression line (default value: true). 
2. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV). 
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it. 

##### NOTE 
- Ordinary least squares regression (OLS) only.  
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV/XLS/XLSX file consisting of two columns.  
- The first column is assumed to be the dependent variable (_y_).  
- Missing data pairs will be removed. 

```{r}
hr()
checkboxInput(inputId = 'confint3', 
              label = 'Plot 95% confidence interval.', 
              value = TRUE)

checkboxInput(inputId = 'type3', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header3', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file3', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
lm_data <- reactive({
    req(input$file3)
    
    if(input$type3 == TRUE) {
        read_csv(input$file3$datapath,
                 col_names = input$header3)
    } else {
        read_excel(input$file3$datapath,
                   col_names = input$header3)
    }
})

renderPrint({
    lm_data()
})
```

### Tabular summary

```{r}
renderPrint({
    skim(lm_data())
})
``` 

Row
-------------------------------------  

### Scatterplot

```{r}
renderPlot({
    if(input$confint3 == TRUE) {
        ggplot(data = lm_data()) +
        aes_string(x = colnames(lm_data())[1],
                   y = colnames(lm_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            geom_smooth(method = 'lm') +
            theme_bw(base_size = 20)
    
    } else {
        ggplot(data = lm_data()) +
        aes_string(x = colnames(lm_data())[1],
                   y = colnames(lm_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
             geom_smooth(method = 'lm',
                         se = FALSE) +
            theme_bw(base_size = 20)
    }
})
```

### Regression summary

```{r}
renderPrint({
    ols <- lm_data() %>% 
        filter(complete.cases(.)) %>% 
        set_names(c('column_1.y', 'column_2.x'))
    
    summary(lm(column_1.y ~ column_2.x, data = ols))
})
```

Contingency tables {data-navmenu='Select a test'}
=====================================  
    
Inputs {.sidebar data-width=350}
-------------------------------------

#### Contingency tables

##### PROCEDURE
1. Select whether you are entering raw data or summarised counts (default value: raw data). 
2. Select whether the uploaded file is in comma-separated values (CSV) format or an MS Excel (XLS or XLSX) format (default value: CSV). 
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it. 

##### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload count data only.  
- If uploading raw data, the file must be a CSV/XLS/XLSX file consisting of two columns, with the data entered in long format.  
- If uploading summarised data, the file must be a CSV/XLS/XLSX file consisting of three columns (two grouping columns and a count column).
- Missing data pairs will be removed. 

```{r}
hr()
selectInput(inputId = 'test4', 
            label = 'Choose a test',
            choices = c("Chi-squared", "Fisher's Exact", 
                        "Mcnemar's"))

selectInput(inputId = 'data_format4', 
            label = 'Choose the data format',
            choices = c("Raw data", "Summarised counts"))

checkboxInput(inputId = 'type4', 
              label = 'File is in CSV format.', 
              value = TRUE)

checkboxInput(inputId = 'header4', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file4', 
          label = 'Upload the file',
          multiple = FALSE,
          accept = c('.csv', '.xls', '.xlsx'))
```

Column
------------------------------------

### Mosaic plot

```{r}
conti_data <- reactive({
    req(input$file4)
    
    if(input$type4 == TRUE) {
        read_csv(input$file4$datapath,
                 col_names = input$header3)
    } else {
        read_excel(input$file4$datapath,
                   col_names = input$header3)
    }
})

renderPlot({
    if(input$data_format4 == 'Raw data') {
        par(cex = 2)
        mosaicplot(table(conti_data()),
                   main = 'Mosaic plot of uploaded data',
                   color = TRUE,
                   cex.axis = 0.75)
    } else {
        tab4 <- conti_data() %>% 
            uncount(count)
        par(cex = 2)
        mosaicplot(table(tab4),
                   main = 'Mosaic plot of uploaded data',
                   color = TRUE,
                   cex.axis = 0.75)
    }
})
```

Column
------------------------------------

### Tabulated data and test results

```{r}
renderPrint({
if(input$test4 == 'Chi-squared') {
    if(input$data_format4 == 'Raw data') {
        CrossTable(x = conti_data()[[1]], 
                   y = conti_data()[[2]],
                   prop.chisq = FALSE,
                   chi = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
    } else {
        tab4 <- conti_data() %>% 
            uncount(count)
        CrossTable(x = tab4[[1]], 
                   y = tab4[[2]],
                   prop.chisq = FALSE,
                   chi = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
        }
    } else if(input$test4 == "Fisher's Exact") {
    if(input$data_format4 == 'Raw data') {
        CrossTable(x = conti_data()[[1]], 
                   y = conti_data()[[2]],
                   prop.chisq = FALSE,
                   fisher = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
    } else {
        tab4 <- conti_data() %>% 
            uncount(count)
        CrossTable(x = tab4[[1]], 
                   y = tab4[[2]],
                   prop.chisq = FALSE,
                   fisher = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
        }
    } else {
    if(input$data_format4 == 'Raw data') {
        CrossTable(x = conti_data()[[1]], 
                   y = conti_data()[[2]],
                   prop.chisq = FALSE,
                   mcnemar = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
    } else {
        tab4 <- conti_data() %>% 
            uncount(count)
        CrossTable(x = tab4[[1]], 
                   y = tab4[[2]],
                   prop.chisq = FALSE,
                   mcnemar = TRUE,
                   dnn = c('Column 1', 'Column 2'),
                   format = 'SPSS')
        }
    }
})
```
