---
title: "StatsAlive"
output: 
  flexdashboard::flex_dashboard:
    navbar:
     - { title: "Developed by PainblogR", icon: 'fa-home', href: "https://www.painblogr.org", align: right }
    social: ['menu']
    source_code: 'https://github.com/kamermanpr/stats-alive.git'
runtime: shiny
---

```{r setup, include = FALSE}
library(flexdashboard)
library(tidyverse)
library(broom)
library(skimr)
skim_with(numeric = list(p0 = NULL, p25 = NULL, p50 = NULL, 
                         p75 = NULL, p100 = NULL, hist = NULL,
                         min = ~ min(., na.rm = TRUE),
                         q25 = ~ quantile(., probs = 0.25, na.rm = TRUE),
                         median = ~ quantile(., probs = 0.5, na.rm = TRUE),
                         q75 = ~ quantile(., probs = 0.75, na.rm = TRUE),
                         max = ~min(., na.rm = TRUE)))

```

Compare two groups
=====================================  


Inputs {.sidebar}
-------------------------------------

#### PROCEDURE
1. Select a hypothesis test (default value: paired Student's _t_-test).  
2. Select whether the first row of the uploaded file includes column headings (default value: true).  
3. Select your data file and upload it.  

#### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV file consisting of two columns.  
- For paired analyses, there must be one column for each of the two measurements made on an animal/person (incomplete pairs will be removed).  
- For unpaired tests, there must be one column for each of the two groups data were collected from.  

```{r}
hr()
selectInput(inputId = 'test', 
            label = 'Choose a test',
            choices = c('Paired: t-test', 'Paired: Wilcoxon signed-rank test', 
                        'Unpaired: t-test', 'Unpaired: Mann-Whitney U test'))

checkboxInput(inputId = 'csv_header', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file_csv', 
          label = 'Upload a CSV file',
          multiple = FALSE,
          accept = ".csv")
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
df <- reactive({
    req(input$file_csv)
    read_csv(input$file_csv$datapath,
             col_names = input$csv_header)
    })

renderPrint({
    df()
})
```

### Summary

```{r}
renderPrint({
    skim(df())
})
``` 

Row
-------------------------------------  

### Box-and-whisker plot

```{r}
renderPlot({
    gathered <- df() %>% 
        gather(key = key,
               value = value)
    
    ggplot(data = gathered) +
        aes(x = key,
            y = value) +
    geom_point(position = position_jitter(width = 0.3)) +
    geom_boxplot(alpha = 0.5) +
    stat_summary(fun.y = 'mean', 
                 geom = 'point',
                 shape = 8,
                 stroke = 1,
                 size = 4) +
    scale_y_continuous(name = 'Value') +
    theme_bw(base_size = 14) +
    theme(axis.title.x = element_blank())
    })
```

### Test summary

```{r}
renderPrint({
    gathered <- df() %>% 
        gather(key = key,
               value = value)
    
    gathered_paired <- df() %>% 
        filter(complete.cases(.)) %>% 
        gather(key = key,
               value = value)
    
    if(input$test == 'Paired: t-test') {
        t.test(value ~ key,
               data = gathered_paired,
               paired = TRUE)
        
    } else if(input$test == 'Unpaired: t-test') {
        t.test(value ~ key,
               data = gathered,
               na.rm = TRUE,
               paired = FALSE)
        
    } else if(input$test == 'Paired: Wilcoxon signed-rank test') {
        wilcox.test(value ~ key, 
                    data = gathered_paired,
                    exact = FALSE,
                    paired = TRUE,
                    confint = TRUE)
        
    } else {
        wilcox.test(value ~ key,
                    data = gathered,
                    na.rm = TRUE,
                    exact = FALSE,
                    paired = FALSE,
                    confint = TRUE)
    }
})
```

Correlation
=====================================  
    
Inputs {.sidebar}
-------------------------------------

#### PROCEDURE
1. Select the type of correlation (default value: Pearson's correlation). 
2. Select whether you want a trend line included on the scatterplot (default value: true).
3. Select whether the first row of the uploaded file includes column headings (default value: true).  
4. Select your data file and upload it. 

#### NOTE 
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV file consisting of two columns.  
- Missing data pairs will be removed. 

```{r}
hr()
selectInput(inputId = 'test2', 
            label = 'Choose a test',
            choices = c("Pearson's correlation", "Spearman's correlation"))

checkboxInput(inputId = 'trend_line2', 
              label = 'Draw a trend line.', 
              value = TRUE)

checkboxInput(inputId = 'csv_header2', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file_csv2', 
          label = 'Upload a CSV file',
          multiple = FALSE,
          accept = ".csv")
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
cor_data <- reactive({
    req(input$file_csv2)
    read_csv(input$file_csv2$datapath,
             col_names = input$csv_header2)
    })

renderPrint({
    cor_data()
})
```

### Summary

```{r}
renderPrint({
    skim(cor_data())
})
``` 

Row
-------------------------------------  

### Scatterplot

```{r}
renderPlot({
    if(input$trend_line2 == TRUE) {
        ggplot(data = cor_data()) +
        aes_string(x = colnames(cor_data())[1],
                   y = colnames(cor_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            geom_smooth(method = 'loess',
                        se = FALSE) +
            theme_bw(base_size = 14)
    
    } else {
        ggplot(data = cor_data()) +
        aes_string(x = colnames(cor_data())[1],
                   y = colnames(cor_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            theme_bw(base_size = 14)
    }
})
```

### Correlation summary

```{r}
renderPrint({
    
    corr <- cor_data() %>% 
        filter(complete.cases(.)) %>% 
        set_names(c('column_1.y', 'column_2.x'))
    
    if(input$test2 == "Pearson's correlation") {
        with(corr, cor.test(x = column_2.x, 
                            y = column_1.y, 
                            method = 'pearson'))
    } else {
        with(corr, cor.test(x = column_2.x, 
                            y = column_1.y, 
                            method = 'spearman',
                            exact = TRUE))
    }
})
```

Regression
=====================================  
    
Inputs {.sidebar}
-------------------------------------

#### PROCEDURE
1. Select whether you want to plot 95% confidence intervals for the regression line (default value: true) . 
2. Select whether the first row of the uploaded file includes column headings (default value: true).
3. Select your data file and upload it.  

#### NOTE 
- Ordinary least squares regression (OLS) only.  
- Results will be generated automatically, and the output updated if you change the selected test.  
- Upload continuous or discrete numeric data only. No count or frequency data.  
- The uploaded file must be a CSV file consisting of two columns.  
- The first column is assumed to be the dependent variable (_y_).  
- Missing data pairs will be removed. 

```{r}
hr()
checkboxInput(inputId = 'confint3', 
              label = 'Plot 95% confidence interval.', 
              value = TRUE)

checkboxInput(inputId = 'csv_header3', 
              label = 'First row is a heading.', 
              value = TRUE)

fileInput(inputId = 'file_csv3', 
          label = 'Upload a CSV file',
          multiple = FALSE,
          accept = ".csv")
```

Row
------------------------------------

### Glimpse at the uploaded data

```{r}
lm_data <- reactive({
    req(input$file_csv3)
    read_csv(input$file_csv3$datapath,
             col_names = input$csv_header3)
    })


renderPrint({
    lm_data()
})
```

### Summary

```{r}
renderPrint({
    skim(lm_data())
})
``` 

Row
-------------------------------------  

### Scatterplot

```{r}
renderPlot({
    if(input$confint3 == TRUE) {
        ggplot(data = lm_data()) +
        aes_string(x = colnames(lm_data())[1],
                   y = colnames(lm_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
            geom_smooth(method = 'lm') +
            theme_bw(base_size = 14)
    
    } else {
        ggplot(data = lm_data()) +
        aes_string(x = colnames(lm_data())[1],
                   y = colnames(lm_data())[2]) +
            geom_point(shape = 21,
                       size = 4,
                       stroke = 0.8,
                       fill = '#FFFFFF',
                       alpha = 0.5) +
             geom_smooth(method = 'lm',
                         se = FALSE) +
            theme_bw(base_size = 14)
    }
})
```

### Regression summary

```{r}
renderPrint({
    ols <- lm_data() %>% 
        filter(complete.cases(.)) %>% 
        set_names(c('column_1.y', 'column_2.x'))
    
    summary(lm(column_1.y ~ column_2.x, data = ols))
})
```

